Monitoring misinformation related interventions by Facebook, Twitter and YouTube: methods and illustration.

### 2021-08-23 article/draft.tex
edit discussion

A last concern regards the growing restrictions to access APIs \cite{api}.
In April 2018, likely as a response to the Cambridge Analytica scandal, Facebook severely limited access to their official APIs
\footnote{\url{https://about.fb.com/news/2018/04/restricting-data-access/}, \url{http://thepoliticsofsystems.net/2018/08/facebooks-app-review-and-how-independent-research-just-got-a-lot-harder/}}, 
and CrowdTangle is today a reliable source for researchers to access Facebook data. 
But CrowdTangle has recently been used by journalists to show that Facebook is spreading biased information
\footnote{\url{https://www.economist.com/graphic-detail/2020/09/10/facebook-offers-a-distorted-view-of-american-news}, \url{https://twitter.com/facebookstop10}. 
The last link is a Twitter account showing each day the ten most shared link on Facebook, most of them coming from extreme right and misinformation sources. 
Facebook has reacted to this here: \url{https://about.fb.com/news/2020/11/what-do-people-actually-see-on-facebook-in-the-us/}}
, and a general suspicion is growing that the access of journalists and researchers to the API might be shut down in a near future.
\footnote{\url{https://www.nytimes.com/2021/07/14/technology/facebook-data.html}, \url{https://www.bloomberg.com/news/articles/2021-08-03/facebook-disables-accounts-tied-to-nyu-research-project}}
Regarding Twitter and YouTube, it is still possible and relatively easy to access their official APIs.%, although in our personal experience we never managed to get access to the Twitter API when we described our intend to use the Twitter developer platform as `academic'.\footnote{The token was given only when we described a `hobbyist' intent.}
It should also be noted that the number of Tweets that can be retrieved has decreased between the Twitter API v1 and v2. 
In the v2, there is a monthly cap of 10,000,000 Tweets for the academic research track
\footnote{https://developer.twitter.com/en/docs/projects/overview},
while there was no monthly limit in the v1, only a limit in the number of requests that could be made in a 15-minute window
\footnote{https://developer.twitter.com/en/docs/twitter-api/v1/rate-limits}.
We thus noticed that we could get around 10 times more Tweets per month with the v1 when compared to the v2 for one specific collection.

@@

{\bf A need for more consistent and adapted policies.}
Both Infowars and The Beauty of Life domains are considered as spreading misinformation (they are both on the Iffy+ list\footnote{https://iffy.news/iffy-plus/}), and violated Facebook's guidelines (incitation to violence and coordinated unauthentic behavior). 
Yet one's distribution was reduced for only half a year, while the other one's links cannot be shared anymore on Facebook since the end of 2019, and it is not clear why such double standard was applied on these comparable websites.
Misinformation on the web is a problem that is increasingly discussed since the 2016 American elections, and in the years that followed, the platforms may have decided manually and on a case-by-case basis which action to enforce.
It is also possible that platforms usually follow specific rules on how to fight misinformation, but that sometimes the sudden mediatization of specific events leads them to take exceptional actions, such as with the Capitol riots and Trump's suspension from Facebook, Twitter and YouTube.

@@

{\color{brown} Recommendation to fight misinformation: inform about sources, example inform users that this account has shared x failed fact-checks over a x-months period.}

@@

%Networked public square / private companies / regulations / coordination...

%User side: indirect effects, strategies... business model

%Make point about how the three platforms communicate differently about their policies. 
%Need for a consistent policy for all, but this will surely come only when a law against misinformation is issued. 
%Facebook's strategy to tackle ``False News'' is three fold : Remove, Reduce, Inform. It is explained via a blog post on the facebook Newsroom. 
%Similarly, Twitter communicates about actions related to misinformation via their \href{https://blog.twitter.com/en\_us/authors.TwitterSafety}{Twitter Safety} Blog/account. 
%In a post on YouTube Official Blog, the platform explained its ``Four Rs of Responsibility" and how it raises authoritative content, reduces borderline content and harmful misinformation (see table \ref{summary} in Appendix \ref{links}).

>>>>

A last concern regards the growing restrictions to access APIs \cite{api}.
In April 2018, likely as a response to the Cambridge Analytica scandal, Facebook severely limited access to their official APIs\footnote{\url{https://about.fb.com/news/2018/04/restricting-data-access/}, \url{http://thepoliticsofsystems.net/2018/08/facebooks-app-review-and-how-independent-research-just-got-a-lot-harder/}}, 
and CrowdTangle is today a reliable source for researchers to access Facebook data. 
However, it should be noted that CrowdTangle has been recently used by journalists to show that Facebook is spreading biased information\footnote{\url{https://www.economist.com/graphic-detail/2020/09/10/facebook-offers-a-distorted-view-of-american-news}, \url{https://twitter.com/facebookstop10}. 
The last link is a Twitter account showing each day the ten most shared link on Facebook, most of them coming from extreme right and misinformation sources. 
Facebook has reacted to this here: \url{https://about.fb.com/news/2020/11/what-do-people-actually-see-on-facebook-in-the-us/}}, 
and a general suspicion is growing that its access might be shut down in a near future to journalists and researchers.\footnote{\url{https://www.nytimes.com/2021/07/14/technology/facebook-data.html}, \url{https://www.bloomberg.com/news/articles/2021-08-03/facebook-disables-accounts-tied-to-nyu-research-project}}
Regarding Twitter and YouTube, it is still possible and relatively easy to access their official APIs, although the number of Tweets that can be retrieved has decreased between the Twitter API v1 and v2. 
In the v2, there is a monthly cap of 10,000,000 Tweets for the academic research track\footnote{https://developer.twitter.com/en/docs/projects/overview},
while there was no monthly limit in the v1, only a limit in the number of requests that could be made in a 15-minute window\footnote{https://developer.twitter.com/en/docs/twitter-api/v1/rate-limits}.
We thus noticed that we could get around 10 times more Tweets per month with the v1 when compared to the v2 for one specific collection.

@@

{\bf A need for more consistent and adapted policies.}
Both Infowars and The Beauty of Life domains are considered as spreading misinformation (they are both on the Iffy+ list\footnote{https://iffy.news/iffy-plus/}), and violated Facebook's guidelines (incitation to violence and coordinated unauthentic behavior). 
Yet one's distribution was reduced for only half a year, while the other one's links cannot be shared anymore on Facebook since the end of 2019, and it is not clear why such double standard was applied on these comparable websites.
Misinformation on the web is a problem that is increasingly discussed since the 2016 American elections, and in the years that followed, the platforms may have decided manually and on a case-by-case basis which action to enforce.
It is also possible that platforms usually follow specific rules on how to fight misinformation, but that sometimes the sudden mediatization of specific events leads them to take exceptional actions, such as with the Capitol riots and Trump's suspension from Facebook, Twitter and YouTube.
We would argue that to efficiently curb the spread of misinformation, policies should be clearly stated and consistently applied, instead of inteventions being implemented on a case-by-case basis and decided in response to the latest news events.

@@

@@

% Recommendation to fight misinformation: inform about sources, example inform users that this account has shared x failed fact-checks over a x-months period.

% Networked public square / private companies / regulations / coordination...

% User side: indirect effects, strategies... business model

% Make point about how the three platforms communicate differently about their policies. 
%Need for a consistent policy for all, but this will surely come only when a law against misinformation is issued. 
%Facebook's strategy to tackle ``False News'' is three fold : Remove, Reduce, Inform. It is explained via a blog post on the facebook Newsroom. 
%Similarly, Twitter communicates about actions related to misinformation via their \href{https://blog.twitter.com/en\_us/authors.TwitterSafety}{Twitter Safety} Blog/account. 
%In a post on YouTube Official Blog, the platform explained its ``Four Rs of Responsibility" and how it raises authoritative content, reduces borderline content and harmful misinformation (see table \ref{summary} in Appendix \ref{links}).
